FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-19 NewVersion/paper.tex
Abstract intro overview
Added all comments (including marcos).
S2 reduction steps revised

%\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
%\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
%\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
%\newcommand\marcoT[1]{\authorText{blue}{#1}}
%\newcommand\haoyuan[1]{\authornote{haoyuan}{orange}{#1}}

\newcommand\bruno[1]{}
\newcommand\yanlin[1]{}
\newcommand\marco[1]{}
\newcommand\haoyuan[1]{}

@@

@@

\begin{abstract}
Multiple inheritance is a valuable feature for Object-Oriented
Programming. However, it is also tricky to get right, as illustrated by
the extensive literature on the topic. A key issue 
is the \emph{ambiguity} arising from inheriting multiple parents,
which can have conflicting methods and fields. 
Numerous existing work provides solutions for 
conflicts which arise from \emph{diamond inheritance}: i.e.
conflicts that arise from implementations sharing a common 
ancestor. However, most mechanisms are inadequate to deal 
with \emph{unintentional method conflicts}: conflicts which 
arise from two unrelated methods that happen to share the same name
and signature.

>>>>

\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
\newcommand\marcoT[1]{\authorText{blue}{#1}}
\newcommand\haoyuan[1]{\authornote{haoyuan}{orange}{#1}}

%\newcommand\bruno[1]{}
%\newcommand\yanlin[1]{}
%\newcommand\marco[1]{}
%\newcommand\haoyuan[1]{}

@@

\lstdefinestyle{reduction}{
  	language=C,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,override},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\sffamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt,
  moredelim=[is][\underbar]{_}{_},
  keepspaces=true
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

@@

\begin{abstract}
Multiple inheritance is a valuable feature for Object-Oriented
Programming. However, it is also tricky to get right, as illustrated by
the extensive literature on the topic. A key issue 
is the \emph{ambiguity} arising from inheriting multiple parents,
which can have conflicting methods. 
Numerous existing work provides solutions for 
conflicts which arise from \emph{diamond inheritance}: i.e.
conflicts that arise from implementations sharing a common 
ancestor. However, most mechanisms are inadequate to deal 
with \emph{unintentional method conflicts}: conflicts which 
arise from two unrelated methods that happen to share the same name
and signature.

### 2017-10-19 NewVersion/sections/Overview.tex
S2 reduction steps revised

\vspace{3pt}\begin{lstlisting}
interface SafeDeck extends Deck {
  boolean isEmpty() {...}
  void draw() { // overriding
    if (isEmpty()) println("The deck is empty.");
    else println("Draw a card");
  }
}
\end{lstlisting}\vspace{3pt}
Without dynamic dispatching, we may have to copy the \lstinline|shuffleAndDraw| code into \lstinline|SafeDeck|, so that \lstinline|shuffleAndDraw| calls the new \lstinline|draw| defined in \lstinline|SafeDeck|. Yet dynamic dispatching immediately saves us from the duplication work,
since the method becomes automatically dispatched to the most specific one. Nevertheless, as seen before, dynamic dispatch would potentially introduce ambiguity.
For instance, when we have the class hierarchy structure shown in Figure~\ref{fig:drawablesafedeck}(left) and the following code:

% \begin{figure*}[t]
% 	\saveSpaceFig
% 	\centering
% 	\includegraphics[height=2cm]{pics/DrawableSafeDeck.pdf}
% 	\caption{UML graph for \lstinline|DrawableSafeDeck|.}\label{fig:drawablesafedeck}
% 	\saveSpaceFig
% \end{figure*}

@@

\vspace{3pt}\begin{lstlisting}
interface DrawableSafeDeck extends Drawable, SafeDeck {}

// main program
new DrawableSafeDeck().shuffleAndDraw()
\end{lstlisting}\vspace{3pt}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck|
is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the
\lstinline|draw| method seems to be ambiguous since
\lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from
both \lstinline|SafeDeck| and \lstinline|Drawable|. 
Indeed, using the a reduction model similar to FJ~\cite{} (where static types tracked), the 
reduction steps would roughly be:
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [new DrawableSafeDeck().shuffle(), new DrawableSafeDeck().draw()]
-> [..., ambiguous call!}]
\end{lstlisting}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck| is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver 
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the \lstinline|draw| method is ambiguous since \lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from both \lstinline|SafeDeck| and \lstinline|Drawable|. But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch.

@@

%%\paragraph{Problem:} \\

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, the method call \lstinline|e.m()|
makes use of both the static type and the dynamic type of the
receiver, so it is a
combination of static and dynamic dispatch. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch. The formal algorithm will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.

@@

Here are the reductions performed in \MIM{} to accept this call: 
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [..., (Deck)new DrawableSafeDeck().draw()]
-> [..., SafeDeck.draw()]
\end{lstlisting}
Notably, we make use of the branch information by adding the cast \lstinline|Deck|, since the type of this-reference (\lstinline|Deck|) specifies the branch to use unambiguously. 
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. This is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned. The details of how this is done will be explained in Section~\ref{sec:formalization}.

>>>>

\vspace{3pt}\begin{lstlisting}
interface SafeDeck extends Deck {
  boolean isEmpty() {...}
  void draw() { // overriding
    if (isEmpty()) println("The deck is empty.");
    else println("Draw a card");
  }
}
\end{lstlisting}\vspace{3pt}
Without dynamic dispatching, we may have to copy the \lstinline|shuffleAndDraw| code into \lstinline|SafeDeck|, so that \lstinline|shuffleAndDraw| calls the new \lstinline|draw| defined in \lstinline|SafeDeck|. Dynamic dispatching immediately saves us from the duplication work,
since the method becomes automatically dispatched to the most specific one. Nevertheless, as seen before, dynamic dispatch would potentially introduce ambiguity.
For instance, when we have the class hierarchy structure shown in Figure~\ref{fig:drawablesafedeck}(left) with the following code:

@@

\vspace{3pt}\begin{lstlisting}
interface DrawableSafeDeck extends Drawable, SafeDeck {}
// main program
new DrawableSafeDeck().shuffleAndDraw()
\end{lstlisting}\vspace{3pt}

\noindent Indeed, using the a reduction model similar to FJ~\cite{Igarashi01FJ} (without static types tracked), the reduction steps would roughly be:

% \begin{equation}\label{}
% \begin{split}
% A & = \frac{\pi r^2}{2} \\
%  & = \frac{1}{2} \pi r^2
% \end{split}
% \end{equation}

\begin{lstlisting}[style=reduction, escapechar=\%]
   %\underline{new DrawableSafeDeck()}%.shuffleAndDraw()
-> %\underline{new DrawableSafeDeck().shuffle()}%, new DrawableSafeDeck().draw()
-> ..., %\underline{new DrawableSafeDeck().draw()}%
-> ..., ambiguous call!!!
\end{lstlisting}

\noindent When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck|
is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the
\lstinline|draw| method seems to be ambiguous since
\lstinline|DrawableSafeDeck| inherits two \lstinline|draw| methods from
both \lstinline|SafeDeck| and \lstinline|Drawable|. 
But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch.

@@

\paragraph{\MIM's solution:} 
The essential problem is how to ensure that the correct method is
invoked. To solve this problem, \MIM{} uses a variant of method
dispatching that we call \textit{hierarchical dispatching}. In
hierarchical dispatching both the static and dynamic type information 
are used to select the right method implementation.
%Inspired by C++, we have formalized this feature into a minimal core calculus,
%studied and proved the type soundness. It is concise and can readily be imported to other systems. 
%And in the formalization, we propose \textit{hierarchical dispatch} in our model
%as the default method lookup algorithm. 
During runtime, a method call \lstinline|e.m()|
makes use of both the static type and the dynamic type of the
receiver, so it is a
combination of static and dynamic dispatching. Intuitively, the static type specifies one branch
to avoid ambiguity, and the dynamic type finds the most specific implementation on that branch. The formal algorithm will be introduced in Section~\ref{sec:formalization} and Section~\ref{sec:auxdefs}.

@@

\noindent Here are the reductions performed in \MIM{}:
\begin{lstlisting}[style=reduction, escapechar=\%]
   %\underline{new DrawableSafeDeck()}%.shuffleAndDraw()
-> %\underline{((DrawableSafeDeck)new DrawableSafeDeck()).shuffleAndDraw()}%
-> %\underline{((Deck)new DrawableSafeDeck()).shuffle()}%, ((Deck)new DrawableSafeDeck()).draw()
-> ..., %\underline{((Deck)new DrawableSafeDeck()).draw()}%
-> ..., ... %//SafeDeck.draw()%
\end{lstlisting}

Notably, we make use of the branch information by adding the cast \lstinline|Deck|, since the type of this-reference (\lstinline|Deck|) specifies the branch to use unambiguously. 
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. This is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned. The details of how this is done will be explained in Section~\ref{sec:formalization}.
